---
title: "Lucas Kantorowski"
format: 
  dashboard:
    orientation: columns
    theme: vapor
    nav-buttons:
      - icon: github
        href: https://github.com/LKantoro/lab-3
        aria-label: GitHub
      - icon: linkedin
        href: https://www.linkedin.com/in/lucas-kantorowski/
        aria-label: LinkedIn
logo: images/wellcomelogo.png
logo-alt: Wellcome Organization Logo
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(gt)
library(dplyr)
library(leaflet)
library(dplyr)
library(sf)
library(plotly)
library(readxl)
```

```{r}
#| label: read-in-wellcome-data

wgm2018_crosstab = read_excel(here::here("data", "wgm2018-dataset-crosstabs-all-countries.xlsx"), sheet = 1)

wgm2018_full = read_excel(here::here("data", "wgm2018-dataset-crosstabs-all-countries.xlsx"), sheet = 2)

wgm2018_dict = read_excel(here::here("data", "wgm2018-dataset-crosstabs-all-countries.xlsx"), sheet = 3)
```


```{r}
#| label: country_label_prep
#creating string based on list provided
countriesstringsplit = str_split(wgm2018_dict[1,3], ',')

#changing the vector into a df and eliminating the last row, which was blank
countriesdf = as.data.frame(countriesstringsplit, col.names = c("country")) %>%
  filter(!row_number() %in% c(145))

#creating a two column data frame about the '=' and changing the code to numeric
countriesdf = countriesdf %>% 
  separate_wider_delim(country, delim = "=", 
                       names = c("country_code", "country_name")) %>% 
  mutate(country_code = as.numeric(country_code))
```


```{r}
#| label: region_label_prep


#similar process for splitting regions, as was done above for countries

regionsstringsplit = str_split(wgm2018_dict[57,3], ',')

regionsdf = as.data.frame(regionsstringsplit, col.names = c("region")) %>%
  filter(!row_number() %in% c(20))

regionsdf = regionsdf %>% 
  separate_wider_delim(region, delim = "=", 
                       names =c("region_code", "region_name")) %>% 
  mutate(region_code = as.numeric(region_code))


#creating a new column for world region to be used to split the graph

regionsdf = regionsdf %>% mutate('map_region' = case_when(
          region_code %in% c(9,10,11,12, 18) ~ 'Asia and Australia',
          region_code %in% c(1,2,4,5) ~ 'Sub-Saharan Africa',
          region_code %in% c(6,7,8) ~ 'Americas',
          region_code %in% c(15,16,17) ~ 'Europe',
          region_code %in% c(3,13) ~ 'Middle East and North Africa',
          region_code == 14 ~ 'Former Soviet Union',
          region_code == 0 ~ 'Not Assigned'))
```


```{r}
#| label: full_data_with_countries_and_regions

wgm2018_full_new1 = left_join(wgm2018_full, countriesdf, 
                              by = join_by(WP5 == country_code))

wgm2018_full_new2 = left_join(wgm2018_full_new1, regionsdf, 
                              by = join_by(Regions_Report == region_code))

countries_and_regions = wgm2018_full_new2 %>% 
  select(country_name, region_name, map_region) %>% distinct()

```



# Science vs. Vaccines

## Column - Plots {width="55%"}


### Row {height="70%"}


```{r}
#| label: plotly_prep_1

#finding the count of surveyed people
country_counts = wgm2018_full_new2 %>% 
  group_by(country_name) %>% 
  summarise(total_count = n()) %>%
  rename('total_surveyed' = total_count)

#finding the count of people who think vaccines are effective
vaccine_effective_counts = wgm2018_full_new2 %>%
  filter(Q26 %in% c('1'), map_region != 'Not Assigned') %>%
  group_by(country_name, map_region) %>% 
  summarise(total_effective = n())
  

#finding the count of people who are knowledgeable about science
know_science_counts = wgm2018_full_new2 %>% 
  filter(Q1 %in% c('1', '2'), map_region != 'Not Assigned') %>%
  group_by(country_name, map_region) %>% 
  summarise(total_science = n()) 

```

```{r}
#| label: plotly_prep_2


#finding percent of people who think vaccines are effective
percent_effective = inner_join(country_counts, vaccine_effective_counts, 
                               by = join_by(country_name == country_name))

percent_effective$percent_eff = 
  percent_effective$total_effective/percent_effective$total_surveyed

#finding percent of people who are knowledgeable about science
percent_science = inner_join(country_counts, know_science_counts, 
                             by = join_by(country_name == country_name))


percent_science$prop_sci = 
  percent_science$total_science/percent_science$total_surveyed


#joining effective counts with safe counts
percent_effective_and_science = left_join(percent_effective, percent_science, 
                                       by = join_by(country_name == country_name))


#rounding to 2 decimal places for cleanliness
percent_effective_and_science = percent_effective_and_science %>% 
  mutate(percent_eff = round(percent_eff, 2), prop_sci = round(prop_sci, 2))



```


:::{.card title="Test Card"}

```{r}
#| label: plotly_graph

#scatterplot of vaccination rates vs science

plot_ly(data = percent_effective_and_science, 
        x = ~prop_sci, y = ~percent_eff, 
        type = "scatter", mode = "markers", text = ~country_name, 
        color = ~map_region.x) %>%
  layout(plot_bgcolor='white',
    xaxis = list(ticksuffix = "%", range = list(0, 1), 
                 title = "'Strong' or 'Some' Knowledge of Science"),
    yaxis = list(ticksuffix="%", range = list(0, 1), 
                 title = ""),
    title = "Belief in Vaccines Being Effective Versus Knowledge of Science") %>% 
  layout(xaxis = list(tickformat = "%"), yaxis = list(tickformat = "%"),
         legend = list(x = 0.85, y = 0.04, bordercolor = 'black',
                       borderwidth = 2, font = list(size = 10)))
```
:::

### Row {height="30%"}

```{r}
#| label: valueboxes_prep

want_to_know_more_science = wgm2018_full_new2 %>% 
  count(Q8) %>% mutate(prop = prop.table(n)) %>% 
  filter(Q8 == '1') %>% 
  select(prop)



want_to_know_more_disease = wgm2018_full_new2 %>% 
  count(Q9) %>% mutate(prop = prop.table(n)) %>% 
  filter(Q9 == '1') %>% 
  select(prop)

disease_part_of_science = wgm2018_full_new2 %>% 
  count(Q3) %>% mutate(prop = prop.table(n)) %>% 
  filter(Q3 == '1') %>% 
  select(prop)

```


```{r}
#| content: valuebox
#| title: "Want To Learn More About Science"
list(
  icon = "lightbulb-fill",
  color = "success",
  value = paste(round(want_to_know_more_science*100,1), "%")
)

```


```{r}
#| content: valuebox
#| title: "Want To Learn More About Disease"
list(
  icon = "virus",
  color = "info",
  value = paste(round(want_to_know_more_disease*100,1), "%")
)

```


```{r}
#| content: valuebox
#| title: "Believe Diseases are a Part of Science"
list(
  icon = "link",
  color = "light",
  value = paste(round(disease_part_of_science*100,1), "%")
)

```



## Column - Medals by year {width="45%"}



### Row - Vaccines {height="70%" .tabset}





```{r}
#| label: dotplot_1

percent_effective %>% ggplot(aes(x = percent_eff, fill = map_region, 
                              color = map_region)) +
  geom_dotplot(binwidth = .01) + 
  facet_grid(rows = vars(map_region), axis.labels = "margins") +
  labs(x = "", y = "", fill = "",
       title = "Belief Vaccines Are Effective By Country and Region",
       subtitle = "People Who 'Agreed' or 'Strongly Agreed'") +
  scale_x_continuous(labels = scales::percent) + xlim(0, 1) +
  geom_text(aes(x = .5, y = .8, label = map_region)) +
  theme_light() + 
  theme(strip.text = element_blank(), legend.position = "none", 
        panel.grid.minor = element_blank(), panel.grid.major.y =element_blank(),
        plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())
```




```{r}
#| label: dotplot_2

percent_science %>% ggplot(aes(x = prop_sci, fill = map_region, 
                              color = map_region)) +
  geom_dotplot(binwidth = .01) + 
  facet_grid(rows = vars(map_region), axis.labels = "margins") +
  labs(x = "", y = "", fill = "",
       title = "Strong Understanding of Science") +
  scale_x_continuous(labels = scales::percent) + xlim(0, 1) +
  geom_text(aes(x = .5, y = .8, label = map_region)) +
  theme_light() + 
  theme(strip.text = element_blank(), legend.position = "none", 
        panel.grid.minor = element_blank(), panel.grid.major.y =element_blank(),
        plot.title.position = "plot", axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())
```



### Row - Tables {height="30%" .tabset}

::: {.card title="Top 30 Vaccination"}


```{r}
#| label: table_1

#table of safe vaccination rates

wgm2018_full_new2 = wgm2018_full_new2 %>% 
  mutate(Q25factor = factor(Q25, levels = c('1', '2', '3', '4', '5', '99'), 
                            labels = c("Strongly Agree", "Agree",
                                       "Neither", "Disagree",
                                       "Strongly Disagree", "No Answer")))


wgm2018_full_new2 %>% filter(Q25 %in% c('1','2','3','4','5')) %>%
  group_by(country_name, Q25) %>% 
  summarize(total_count = n()) %>% 
  pivot_wider(names_from = 'Q25', values_from = 'total_count') %>%
  slice_head(n = 30) %>%
  arrange(desc('2'))



#%>%
#  gt() %>%
#  cols_align(align = "left", columns = "country_name") %>%
#  data_color(
#    method = "numeric",
#    palette = "nord::aurora"
#  )



#olympics |>
#  count(team, medal) |>
#  pivot_wider(names_from = medal, values_from = n, values_fill = 0) |>
#  mutate(Total = Bronze + Gold + Silver) |>
#  relocate(Team = team, Gold, Silver, Bronze, Total) |>
#  arrange(desc(Total), Team) |>
#  slice_head(n = 30) |>
#  gt() |>
#  cols_align(align = "left", column = "Team") |>
#  data_color(
#    method = "numeric",
#    palette = "nord::aurora"
#  )
```

:::

::: {.card title="Bottom 30 Vaccination"}


```{r}
#| label: table_2

#table of safe vaccination rates

wgm2018_full_new2 %>% 
  group_by(country_name, Q25) %>% 
  summarize(total_count = n()) %>% 
  pivot_wider(names_from = 'Q25', values_from = 'total_count') %>%
  rename("Strongly_Agree" = '1', "Somewhat_Agree" = '2', 
         "Neither" = '3', "Somewhat_Disagree" = '4',
         "Strongly_Disagree" = '5') %>% 
  arrange(Strongly_Agree) %>%
  gt() %>%
  cols_align(align = "left", columns = country_name) %>%
  data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```

:::

# Trust in Science {orientation="columns"}


## Column - Plots {width="55%"}

```{r}
#| label: chloropleth_prep_1

#finding country geometries: https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html

download.file(
  "https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/world_shape_file.zip",
  destfile = "data/world_shape_file.zip"
)

#system("unzip DATA/world_shape_file.zip")

world_sf <- read_sf(paste0(
  getwd(), "/data/world_shape_file/",
  "TM_WORLD_BORDERS_SIMPL-0.3.shp"
))
```

```{r}
#| label: chloropleth_prep_2


#finding the count of people who are knowledgeable about science
trust_scientists_counts = wgm2018_full_new2 %>% 
  filter(Q11C %in% c('1', '2')) %>%
  group_by(country_name, map_region) %>% 
  summarise(total_trust = n()) 




#finding percent of people who are knowledgeable about science
percent_trust = inner_join(country_counts, trust_scientists_counts, 
                             by = join_by(country_name == country_name))


percent_trust$perc_trust = 
  round(percent_trust$total_trust/percent_trust$total_surveyed * 100,2)


#joining country geometries and overall data frame together
chloroplethdf = right_join(percent_trust, world_sf, 
                           by = join_by(country_name == NAME))

```




```{r}
#| label: chloropleth_prep_3


#creating bins for coloring
bins = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

pal = colorBin("Paired", domain = chloroplethdf$perc_trust, bins = bins)

#converting data.frame to sf for graphing using leaflet: https://r-spatial.github.io/sf/reference/sf.html
chloroplethdf = st_sf(chloroplethdf)
```

```{r}
#| label: chloropleth

leaflet(chloroplethdf) %>% addTiles() %>% 
  addPolygons(fillColor = ~pal(perc_trust), weight = .2,  opacity = 1, 
              color = 'black', 
              popup = ~paste(country_name, ":",perc_trust, '%'), 
              label = ~paste(country_name, ":",perc_trust, '%'), 
              stroke = FALSE) %>%
  addLegend(pal = pal, values = ~perc_trust, position = "bottomleft", 
            title = "", labFormat = labelFormat(suffix = "%")) %>%
  addControl(c("Percentage of Population Who Have Trust In Scientists"), 
             position = "topleft") %>%
  setView(lng = 0, lat = 0, zoom = 1.5) 
```

## Column - Medals by year {width="45%"}

### Row - Valueboxes {height="30%"}

```{r}
#| label: valuebox_prep

#Q11C, trust in scientists

#percent of religious people who trust scientists
religious_trust_science = wgm2018_full_new2 %>% filter(D1 == '1') %>%
  count(Q12) %>% mutate(prop = prop.table(n)) %>%
  filter(Q12 == '1') %>% select(prop)


#percent of non-religious people who trust scientists
non_religious_trust_science = wgm2018_full_new2 %>% filter(D1 == '2') %>%
  count(Q12) %>% mutate(prop = prop.table(n)) %>%
  filter(Q12 == '1') %>% select(prop)

```


::: {.valuebox icon="yin-yang" color="#FF0000"}

`{r} paste(round(religious_trust_science*100,1), "%")`

of Religious People Have High Trust in Science

:::


::: {.valuebox icon="slash-circle" color="#000000"}

`{r} paste(round(non_religious_trust_science*100,1), "%")`

of Non-Religious People Have High Trust in Science

:::



### Row - Medals By Year {height="70%"}

```{r}
#| label: barplot

wgm2018_full_new2 %>% filter(Q12 %in% c(1,2,3,4)) %>% 
  filter(Education %in% c(1,2,3)) %>%
  ggplot(aes(x = factor(Education, levels = c('1', '2', '3'), 
                        labels = c('Primary','Secondary', 'Tertiary')),
             fill = factor(Q12, levels = c('4', '3', '2', '1'), 
                           labels = c('None', 'Little', 'Some', 'Strong')))) + 
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = 'Education Level', y = '', fill = 'Belief in Science', 
  title = 'Distribution of Science Belief By Education Level') +
  theme(legend.position = "right", panel.grid = element_blank(),
        axis.ticks.x = element_blank(), axis.ticks.y = element_blank(),
        plot.title.position = "plot", panel.grid.major = element_blank()) +
  scale_fill_manual(values = c('red', 'orange', 'yellow', 'green'))

```






